\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-07-31 paper.qmd
Saved the pbmc, tsne projections with the new function
Update pbmc tsne model

@@

## Remove an edge that exist as long edge in p-D
# tr_from_to_df_pbmc |> mutate(ID = row_number()) |> filter(from == 37) |> filter(to == 43) 
tr_from_to_df_pbmc <- tr_from_to_df_pbmc |>
  filter(row_number() != 360)

@@

<!--bin1 = 15, bin2 = 20, b = 300, non_empty = 136-->
```{r}
#| echo: false
#| fig-cap: "(a) Model generated in $2\\text{-}D$ with tSNE, and (b) $p\\text{-}D$ model error in $2\\text{-}D$. The $2\\text{-}D$ model shows three well-separated distant clusters. The $p\\text{-}D$ model errors are distributed along clusters, but most low $p\\text{-}D$ model errors present in the large cluster."
#| label: fig-model-pbmc
#| fig-pos: H
#| out-height: 30%

::: {#fig-pbmc2-sc layout-ncol="4" fig-pos="H"}
![](figures/pbmc3k/tsne_trimesh_plot.png)

>>>>

tr_from_to_df_pbmc <- tr_from_to_df_pbmc |>
  filter(row_number() != 360)

@@

## Remove an edge that exist as long edge in p-D
# tr_from_to_df_pbmc |> mutate(ID = row_number()) |> filter(from == 37) |> filter(to == 43)

trimesh_removed_pbmc <- vis_rmlg_mesh(
  distance_edges = distance_pbmc,
  benchmark_value = benchmark_pbmc,
  tr_coord_df = tr_from_to_df_pbmc,
  distance_col = "distance") +
  geom_point(
    data = tsne_pbmc_scaled,
    aes(
      x = tSNE1,
      y = tSNE2
    ),
    alpha = 0.3,
    size = 0.1,
    color = "#6a3d9a"
  ) +
  theme_void() #+
  #xlim(sc_xlims) + ylim(sc_ylims) +
  #interior_annotation("a", sc_ltr_pos)

@@

```{r}
#| label: proj-tsne-pbmc

## Hexagonal binning to have regular hexagons
hb_obj_pbmc <- hex_binning(
  data = tsne_pbmc_scaled,
  bin1 = num_bins_x_pbmc,
  r2 = r2_pbmc,
  q = 0.05)

tsne_data_with_hb_id <- hb_obj_pbmc$data_hb_id

df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID),
                                tsne_data_with_hb_id)

### Define type column
df <- df_all_pbmc |>
  dplyr::select(tidyselect::starts_with("PC")) |>
  dplyr::mutate(type = "data") ## original dataset

df_b <- df_bin_pbmc |>
  dplyr::filter(hb_id %in% df_bin_centroids_pbmc$hexID) |>
  dplyr::mutate(type = "model") ## Data with summarized mean

## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
df_b <- df_b[match(df_bin_centroids_pbmc$hexID, df_b$hb_id),] |>
  dplyr::select(-hb_id)

df_exe <- dplyr::bind_rows(df_b, df)

## Set the maximum difference as the criteria
distance_df_small_edges <- distance_pbmc |>
  dplyr::filter(distance < benchmark_pbmc)
## Since erase brushing is considerd.

## First projection
projection <- cbind(
  c(0.038104,-0.001148,0.003905,-0.003332,-0.011747,-0.002688,0.003468,-0.014321,-0.000320),
  c(-0.000902,0.041414,0.005089,-0.000149,-0.000551,0.000239,-0.008013,-0.005799,-0.001411))

proj1_pbmc_tsne <- gen_proj_langevitour(
  points_df = df_exe,
  projection = projection,
  edge_df = distance_df_small_edges |> select(-distance)
) + 
  theme_void() + 
  theme(legend.position = "none")

## Second projection
projection <- cbind(
  c(0.013182,-0.022240,-0.001729,0.026946,-0.005289,0.001235,-0.007784,0.017500,-0.006949),
  c(0.027138,0.011939,0.009374,-0.004582,-0.004771,-0.027107,-0.005287,-0.002325,-0.007951))

proj2_pbmc_tsne <- gen_proj_langevitour(
  points_df = df_exe,
  projection = projection,
  edge_df = distance_df_small_edges |> select(-distance)
) + 
  theme_void() + 
  theme(legend.position = "none")

## Third projection
projection <- cbind(
  c(-0.003540,0.025008,0.023261,-0.012782,-0.011617,-0.006796,0.009566,-0.015022,0.001164),
  c(0.038201,-0.007751,0.005064,-0.014143,0.003255,-0.001091,0.001611,-0.003674,-0.008314))

proj3_pbmc_tsne <- gen_proj_langevitour(
  points_df = df_exe,
  projection = projection,
  edge_df = distance_df_small_edges |> select(-distance)
) + 
  theme_void() + 
  theme(legend.position = "none")

<!-- <!--bin1 = 15, bin2 = 20, b = 300, non_empty = 136--> 
<!-- ```{r} -->
<!-- #| echo: false -->
<!-- #| fig-cap: "(a) Model generated in $2\\text{-}D$ with tSNE, and (b) $p\\text{-}D$ model error in $2\\text{-}D$. The $2\\text{-}D$ model shows three well-separated distant clusters. The $p\\text{-}D$ model errors are distributed along clusters, but most low $p\\text{-}D$ model errors present in the large cluster." -->
<!-- #| label: fig-model-pbmc -->
<!-- #| fig-pos: H -->
<!-- #| out-height: 30% -->

<!-- trimesh_removed_pbmc + error_plot_pbmc +  -->
<!--   plot_layout(guides='collect', ncol=2) & -->
<!--   theme(legend.position='bottom') -->
<!-- ``` -->

```{r}
#| echo: false
#| fig-cap: "Model in $2\\text{-}D$, on the tSNE layout, and three views of the fit in projections from $9\text{-}D$, for the PBMC3k data ($(s_1, \ s_2) = (-0.050, \ -0.058)$, $b = 300 \  (15, \ 20)$, $m = 136$, benchmark value to remove large edges is $0.133$). (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/5Y1hE4i7N2k>)."
#| label: fig-pbmc2-sc
#| fig-pos: H
#| out-height: 100%
#| eval: false

trimesh_removed_pbmc + proj1_pbmc_tsne + 
  proj2_pbmc_tsne + proj3_pbmc_tsne +
  plot_layout(guides='collect', ncol=2) &
  theme(legend.position='none')
```

::: {#fig-pbmc2-sc layout-ncol="2" fig-pos="H"}
![](figures/pbmc3k/tsne_trimesh_plot.png)

### 2024-07-31 paper.tex
Saved the pbmc, tsne projections with the new function
Update pbmc tsne model

\includegraphics[width=1\textwidth,height=0.3\textheight]{paper_files/figure-pdf/fig-model-pbmc-1.pdf}

\caption{\label{fig-model-pbmc}(a) Model generated in \(2\text{-}D\)
with tSNE, and (b) \(p\text{-}D\) model error in \(2\text{-}D\). The
\(2\text{-}D\) model shows three well-separated distant clusters. The
\(p\text{-}D\) model errors are distributed along clusters, but most low
\(p\text{-}D\) model errors present in the large cluster.}

\begin{minipage}{0.25\linewidth}
\includegraphics{figures/pbmc3k/tsne_trimesh_plot.png}\end{minipage}%
%
\begin{minipage}{0.25\linewidth}
\includegraphics{figures/pbmc3k/sc_4.png}\end{minipage}%
%
\begin{minipage}{0.25\linewidth}
\includegraphics{figures/pbmc3k/sc_5.png}\end{minipage}%
%
\begin{minipage}{0.25\linewidth}
\includegraphics{figures/pbmc3k/sc_6.png}\end{minipage}%

>>>>

\begin{minipage}{0.50\linewidth}
\includegraphics{figures/pbmc3k/tsne_trimesh_plot.png}\end{minipage}%
%
\begin{minipage}{0.50\linewidth}
\includegraphics{figures/pbmc3k/sc_4.png}\end{minipage}%
\newline
\begin{minipage}{0.50\linewidth}
\includegraphics{figures/pbmc3k/sc_5.png}\end{minipage}%
%
\begin{minipage}{0.50\linewidth}
\includegraphics{figures/pbmc3k/sc_6.png}\end{minipage}%
